Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-05-16 WholePlotTrajectories.Rmd
Discussion

both the lag between taxonomic and functional recovery and the much better consistency of the IDH when the functional diversity is considered rely on the central role of functional redundancy.
The functional redundancy that is the overlap between species, typical of the huge biodiversity of tropical forests [@Bellwood2006] and determinant for communities resilience [@Trenbath1999; @Elmqvist2003; @Diaz2005], was not recovered 30 years after disturbance.

Disturbance globally reduced the overall functional redundancy of communities but this restored rapidly without consistent rtajectories along time.
The initial functional space corresponding to undistrubed 
Besides the long-term alteration of functional redundancy, there was probably persistent compositional changes favoring disturbance resistant species [@Haddad2008], lianas or epiphytes [@Martin2013] and environmental changes, like in the soils nutrient cycling and compaction [@Olander2005].
These persistent changes highly question forest's resilience [@Chazdon2003a].
New conditions would not only be longer lasting but self-maintained as tied to disturbance regime [@Burslem2000].
Specificaly, this would impair species contingent to undisturbed forests, threatening their maintenance, and run the risk to loose cornerstone species and trigger unexpected ecological consequences [@Jones1994;@Diaz2005;@Gardner2007].

>>>>

Both the lag between taxonomic and functional recovery and the middling consistency of the IDH regarding communities taxonomic diversity entail a central role of functional redundancy, *i.e.* the overlap between species functional spaces.
Functional redundancy is typical of the huge biodiversity of tropical forests [@Bellwood2006] and is determinant of communities resilience [@Trenbath1999; @Elmqvist2003; @Diaz2005], and it seemed not to have recovered 30 years after disturbance.

Although disturbance would be mitigated by communities functional redundancy, we expect the redundancy itself to be reduced or re-organized within the functional space.
Globally, disturbed communities did not follow consistent trajectories after disturbance and 30 years after disturbance, irrespective of the intensity, disturbed communities displayed lower, similar or higher functional redundancy.
The redundancy within the functional space of the initial community however clearly followed humped-back trajectories which decrease was determined by the disturbance intensity and which return to initial state was not observed for any disturbed community.
Not only did the functional redundancy and hence the resilience of pre-disturbance community decrease, but also did the ones of disturbance-specific communities increase.
It means higher chances to have long lasting or self-maintained compositional changes towards disturbance resistant species, lianas or epiphytes [@Haddad2008; @Burslem2000; @Martin2013] and thus highly question forest's resilience [@Chazdon2003a].
Specificaly, this would impair species contingent to undisturbed forests, threatening their maintenance, and run the risk to loose cornerstone species and trigger unexpected ecological consequences [@Jones1994;@Diaz2005;@Gardner2007].
